\section{Methodology}
\label{sec:method}
\para{Problem setup.} For each question $x$ with gold label $y$, we generate answers under four conditions: baseline ($a^{(0)}$), re-decode control ($a^{(r)}$), intrinsic self-critique revision ($a^{(s)}$), and external-critique revision ($a^{(e)}$) on a subset. We evaluate both behavioral change and internal representation drift relative to baseline.

\para{Datasets and sampling.} We use 24 sampled examples from GSM8K test and 24 sampled examples from CommonsenseQA validation (seed 42), yielding 48 total examples with balanced task composition. GSM8K scoring parses numeric final answers; CommonsenseQA scoring parses final option letters A--E.

\para{Models and decoding.} The local model is \qwen (1.5B Instruct). External critique uses \gptfourone for 16 examples due to runtime and API cost constraints. Decoding uses temperature 0.0 and max 96 new tokens.

\para{Representation extraction.} Let $\vh^{(c)}_{i,\ell}$ be the final-token residual representation for example $i$, layer $\ell$, and condition $c\in\{r,s,e\}$. We compute cosine drift from baseline
\begin{equation}
D^{\cos}_{i,\ell}(c)=1-\frac{\vh^{(0)}_{i,\ell}\cdot\vh^{(c)}_{i,\ell}}{\|\vh^{(0)}_{i,\ell}\|\,\|\vh^{(c)}_{i,\ell}\|},
\end{equation}
and L2 drift $D^{\ell_2}_{i,\ell}(c)=\|\vh^{(c)}_{i,\ell}-\vh^{(0)}_{i,\ell}\|_2$.

\para{Statistical analysis.} We report accuracy deltas with bootstrap 95\% CIs, McNemar tests for paired correctness changes, and paired Wilcoxon signed-rank tests for layer-wise drift differences between self-critique and re-decode, corrected with Benjamini--Hochberg FDR across 28 layers.

\para{Probe analysis.} We train logistic-regression probes on drift profiles to classify whether initially incorrect examples are corrected after self-critique. Because only 2 of 29 probe examples are positive, we treat probe metrics as diagnostic rather than confirmatory.

\para{Implementation details.} We run Python 3.12.8 with torch 2.10.0+cu128, transformers 5.3.0, datasets 4.6.1, scikit-learn 1.8.0, scipy 1.17.1, and statsmodels 0.14.6. Experiments were executed on two RTX 3090 GPUs with mixed-precision autocast and sequential generation, taking about 13 minutes for the main 48-example run.
